---
title: "data-wrangling_codes"
output: html_document
date: "2023-08-16"
---

# Data wrangling

## Typing task

- Remove all the fast (<0.25s) and slow (> 5s) trials

```{r}
library(tidyverse)

typing_df <- read_csv("consolidated_data/consolidated_typing_data.csv", 
                      col_types = cols(group = col_character(),
                                       typedNumber = col_character(),
                                       typingRT = col_double(),
                                       answerCorr = col_double(),
                                       participant = col_character())) %>%
  filter(typingRT >= 0.25 & typingRT <= 5) %>% 
  group_by(participant) %>%
  summarize(median_RT = median(typingRT),
            mean_RT = mean(typingRT),
            num_correct_trials = sum(answerCorr == 1, na.rm = TRUE),
            perc_accuracy = num_correct_trials / 28 * 100) 
```


```{r}
# Save the cleaned data
write_csv(typing_df, "consolidated_data/cleaned_typing_data.csv")
  
```

## Multiplication production task

```{r}
# Define column types based on the message
col_types <- cols(
  problem = col_factor(),
  OperandType = col_factor(),
  productType = col_factor(),
  group = col_character(),
  typedAnswer = col_character(),
  multiplyRT = col_double(),
  multiplyCorr = col_double(),
  participant = col_character()
)

mult_ACC <- read_csv("consolidated_data/consolidated_mult_data.csv", 
                    col_types = col_types) %>% 
  filter(multiplyRT <= 10) %>% # remove extreme RT
  filter(multiplyRT >= 0.25) 

mult_RT <- read_csv("consolidated_data/consolidated_mult_data.csv", 
                    col_types = col_types) %>% 
  filter(multiplyCorr == 1) %>% 
  filter(multiplyRT <= 10) %>% # remove extreme RT
  filter(multiplyRT >= 0.25) 

# Calculate the trials for each participant
n_trials_mult <- mult_ACC %>% 
  group_by(participant) %>% 
  summarize(n = n()) 

rt_extreme <- mult_ACC %>% 
  filter(multiplyRT <= 10) %>% 
  filter(multiplyRT >= 0.25)

stats_accuracy <- mult_ACC %>% 
  filter(multiplyCorr == 1) %>% 
  group_by(participant) %>% 
  summarize(n = n(),
            mean_mult_RT = mean(multiplyRT),
            medium_mult_RT = median(multiplyRT),
            perc_accuracy_mult = n()/60*100) 
```


```{r}
# Save the cleaned data
write_csv(mult_ACC, "consolidated_data/cleaned_mult-acc_data.csv")
write_csv(stats_accuracy, "consolidated_data/cleaned_mult-summary.csv")
write_csv(mult_RT, "consolidated_data/cleaned_mult_data.csv")
write_csv(mult_RT, "consolidated_data/cleaned_mult-RT_data.csv")  
```


## Digit span task

```{r}
col_types <- cols(
  digitSpan = col_character(),
  digits_audio = col_character(),
  EgnerDSrt = col_double(),
  EgnerDSscore = col_double(),
  group = col_character(),
  participant = col_character()
)

egner_span_df <- read_csv("consolidated_data/consolidated_digit-span-egner_data.csv", 
                    col_types = col_types) %>% 
  filter(EgnerDSscore == 1) %>% 
  filter(!is.na(digitSpan)) %>% 
  group_by(participant) %>%
  summarize(max_span = max(digitSpan)) %>% 
  mutate(digitspan_type = "Egner") 
```


```{r}
# Save the cleaned data
write_csv(egner_span_df, "consolidated_data/cleaned_digit-span-egner_data.csv") 
```


```{r}
col_types <- cols(
  digitSpanWAIS = col_character(),
  WAISDSresponse = col_character(),
  WAISDSrt = col_double(),
  WAISDSscore = col_double(),
  group = col_character(),
  participant = col_character()
)

wais_span_df <- read_csv("consolidated_data/consolidated_digit-span-wais_data.csv", 
                    col_types = col_types) %>% 
  filter(WAISDSscore == 1) %>% 
  filter(!is.na(digitSpanWAIS)) %>% 
  group_by(participant) %>%
  summarize(max_span = max(digitSpanWAIS)) %>% 
  mutate(digitspan_type = "WAIS") 
```


```{r}
# Save the cleaned data
write_csv(wais_span_df, "consolidated_data/cleaned_digit-span-wais_data.csv") 
```

# Combined digit span tasks in two format (wide and long)

```{r}
# Combined digit span tasks (long)

combined_ds <- bind_rows(wais_span_df, egner_span_df) %>% 
    mutate(max_span = as.numeric(max_span)) 
```


```{r}
write_csv(combined_ds, "consolidated_data/cleaned_digit-span-combined_long_data.csv")
```


```{r}
# Combined digit span tasks (wide)
combined_ds <- bind_rows(wais_span_df, egner_span_df) %>% 
    mutate(max_span = as.numeric(max_span))  %>% 
  pivot_wider(
    names_from = digitspan_type,
    values_from = max_span,
    names_prefix = "verbal_span_"
  ) %>% 
    mutate(verbal_span_wais = as.numeric(verbal_span_WAIS),  # Ensure columns are treated as numeric
         verbal_span_egner = as.numeric(verbal_span_Egner)) %>%
    dplyr::select(participant, verbal_span_wais, verbal_span_egner) %>% 
  mutate(verbal_span_mean = (verbal_span_wais + verbal_span_egner)/2)
```


```{r}
write_csv(combined_ds, "consolidated_data/cleaned_digit-span-combined_wide_data.csv")
```


## Visual Pattern task

```{r}
col_types <- cols(
  vptCorrect = col_character(),
  noOfBlocks = col_character(),
  group = col_character(),
  participant = col_character()
)

vpt_df <- read_csv("consolidated_data/consolidated_vpt_data.csv", 
                    col_types = col_types) %>% 
  filter(vptCorrect == "TRUE") %>%
  group_by(participant) %>%
  slice_tail(n = 3) %>% 
  summarize(mean_blocks = mean(as.numeric(noOfBlocks)))

# Highest block span (two out of three correct trials in each block)
vpt_max <- read_csv("consolidated_data/consolidated_vpt_data.csv", 
                   col_types = col_types) %>%
  filter(vptCorrect == "TRUE") %>%
  group_by(participant, noOfBlocks) %>%
  filter(sum(vptCorrect == "TRUE") >= 2) %>%
  ungroup() %>% 
  group_by(participant) %>%
  summarise(max_vpt_span = max(noOfBlocks))

vpt_df <- vpt_df %>%
  left_join(vpt_max, by = "participant") 
```


```{r}
# Save the cleaned data
write_csv(vpt_df, "consolidated_data/cleaned_vpt_data.csv") 
```

## Demographics data for all participants

```{r}

demo_df <- read_csv("consolidated_data/combined_demo_data.csv") %>% 
  mutate(handedness = recode(handedness,
    "Right-handed" = "Right-handed",
    "Right" = "Right-handed",
    "Left-handed" = "Left-handed",
    "Left" = "Left-handed"
  )) %>% 
  dplyr::select(-dob)

```

## Task dataset for all participants

```{r}

typing_df <- read_csv("consolidated_data/cleaned_typing_data.csv")
mult_summary_df <- read_csv("consolidated_data/cleaned_mult-summary.csv")
verbal_stm_df <- read_csv("consolidated_data/cleaned_digit-span-combined_wide_data.csv") 
visual_stm_df <- read_csv("consolidated_data/cleaned_vpt_data.csv") 

```

# Combined dataset

```{r}
combined_tasks <- typing_df %>% 
  inner_join(verbal_stm_df, by = 'participant') %>% 
  inner_join(visual_stm_df, by = 'participant') %>% 
  inner_join(mult_summary_df, by = 'participant') %>% 
  filter(!is.na(verbal_span_mean))

# Check for missing/NA cells in the entire data frame
any_missing <- any(is.na(combined_tasks))

# Check for missing/NA cells for each participant individually
participants_with_missing <- combined_tasks %>%
  group_by(participant) %>%
  summarise(has_missing = any(is.na(.)))
```


```{r}
# save the combined data
write_csv(combined_tasks, "combined_tasks.csv") 
```
Removing na in typing task reduces the n of participant from 240 -236
removing na in verbal span task (digit span) reduces the participant from 236 to 232.

Final sample is 232 (removed 8 participants for incomplete tasks)


