---
title: "EDA"
output: html_document
date: "2023-09-19"
---

## Demographics data for participants

```{r}

# Demographics
library(tidyverse)
demo_df <- read_csv("consolidated_data/combined_demo_data.csv") %>% 
  mutate(handedness = recode(handedness,
    "Right-handed" = "Right-handed",
    "Right" = "Right-handed",
    "Left-handed" = "Left-handed",
    "Left" = "Left-handed"
  ))



# Behavioral Tasks
typing_df <- read_csv("consolidated_data/cleaned_typing_data.csv")
mult_summary_df <- read_csv("consolidated_data/cleaned_mult-summary.csv")
verbal_stm_df <- read_csv("consolidated_data/cleaned_digit-span-combined_wide_data.csv") %>% 
  dplyr::select(participant, verbal_span_mean)
visual_stm_df <- read_csv("consolidated_data/cleaned_vpt_data.csv") 

```


# Combined all the tasks together in a single dataset

```{r}
combined_df <- typing_df %>% 
  left_join(verbal_stm_df, by = 'participant') %>% 
  left_join(visual_stm_df, by = 'participant') %>% 
  left_join(mult_summary_df, by = 'participant') %>% 
  filter(!is.na(median_RT)) %>% # filter the NA in typing task
  filter(!is.na(mean_blocks)) %>% 
  filter(!is.na(mean_mult_RT)) %>% 
  filter(!is.na(verbal_span_mean)) 

#write_csv(combined_df, "combined_data.csv")

# Filter participants in combined_df that are also in combined_tasks
combined_df <- combined_df %>%
  filter(participant %in% combined_tasks$participant)

```

Note for exclusion.
Participant 987 did not complete verbal - digit span task
participant 809 did not complete verbal - digit span task

Removing na in typing task reduces the n of participant from 240 -236
removing na in verbal span task (digit span) reduces the participant from 236 to 231.

Final sample is 231 (removed 9 participants for incomplete tasks)

```{r}
  
# Final count now is 
nrow(combined_df)
```

## Information for demographic table.
### Age

```{r}

# Count the total number of participants
total_participants <- nrow(combined_df)

# Count the number of females
female_count <- sum(!is.na(combined_df$gender) & combined_df$gender == "Female")

# Calculate the percentage of females
female_percentage <- (female_count / total_participants) * 100

# Calculate the age statistics
mean_age <- mean(combined_df$age, na.rm = TRUE)
sd_age <- sd(combined_df$age, na.rm = TRUE)


# Print the statement
cat("The final sample consisted of", total_participants, "young Malaysians (",
    female_count, "females, ", sprintf("%.2f%%", female_percentage), 
    ") aged", min_age <- min(combined_df$age, na.rm = TRUE), "to", max(combined_df$age, na.rm = TRUE), 
    "years (M =", mean_age, ", SD =", sd_age, ").\n")

```

# Task performance (main)
```{r}
library(pastecs)

descriptive_table <- stat.desc(combined_df, norm = TRUE)

write_csv(descriptive_table, "EDA_output/task_performance_metrics.csv")

```


# Rate Correct Score (RCS)

Reaction times and accuracy were used to calculate the rate-correct score (RCS) (Woltz and Was, 2006), a measure of efficiency. RCS is calculated by the number of correct trials divided by the sum of reaction time for correct and incorrect trials, providing thus a measure that combines accuracy and reaction times. The value of RCS indicates the number of correct trials per second, where a higher value of RCS denotes higher efficiency. RCS has been shown to be more efficient in effect detection and accounting for a larger proportion of the variance compared to other integrative measures of speed and accuracy (Vandierendonck, 2017).

```{r}

```



## RCS for typing task

```{r}

```

## RCS for multiplication task

```{r}

```

# Correlation

```{r}

combined_corr_df <- combined_df %>% 
  dplyr::select(age,
    median_RT, mean_RT,
    num_correct_trials, perc_accuracy,
    verbal_span_mean, mean_blocks,
    medium_mult_RT, mean_mult_RT,
    perc_accuracy_mult
  )
```

## Correlation matrix

```{r}
library("Hmisc")
library("corrplot")

# Impute missing values in the "age" column with the mean
combined_corr_df$age[is.na(combined_corr_df$age)] <- mean(combined_corr_df$age, na.rm = TRUE)

# Calculate correlation coefficients and p-values
mydata.rcorr <- rcorr(as.matrix(combined_corr_df))

# Extract coefficient matrix and p-value matrix
correlation_matrix <- mydata.rcorr$r
p_value_matrix <- mydata.rcorr$P

# Calculate correlation coefficients
correlation_matrix <- cor(as.matrix(combined_corr_df), method = "spearman")

# Save the correlation matrix to a CSV file
write.csv(correlation_matrix, "EDA_output/correlation_matrix.csv")

# Save the p-value matrix to a CSV file
write.csv(p_value_matrix, "EDA_output/p_value_matrix.csv")


```


