---
title: "DESCRIPTIVE-stats-code"
output: html_document
date: "2023-08-16"
---

# Import libraries
```{r}
library(tidyverse)
library(janitor)

# Load necessary packages
library(dplyr)
library(psych)

library(apaTables)
library(openxlsx)

```

# Data frames

```{r}

demo_df <- read_csv("consolidated_data/combined_demo_data.csv", col_types = cols(
  participant = col_character(),  # Set 'participant' as character type
  gender = col_character(),
  nationality = col_character(),
  handedness = col_character(),
  age = col_double(),
  dob = col_date(format = "%Y-%m-%d")  
))


mult_df <- readr::read_csv("consolidated_data/cleaned_mult-summary.csv", col_types = cols(
  participant = col_character(),  # Set 'participant' as character type
  n = col_double(),
  mean_mult_RT = col_double(),
  medium_mult_RT = col_double(),
  perc_accuracy_mult = col_double()
))


vpt_df <- readr::read_csv("consolidated_data/cleaned_vpt_data.csv", col_types = cols(
  participant = col_character()  # Set 'participant' as character type
)) %>% 
  select(-max_vpt_span)  # Remove the 'max_vpt_span' column

digit_df <- readr::read_csv("consolidated_data/cleaned_digit-span-combined_data.csv", col_types = cols(
  participant = col_character(),  # Set 'participant' as character type
  max_span = col_double(),
  digitspan_type = col_character()
))

typing_df <- readr::read_csv("consolidated_data/cleaned_typing_data.csv", col_types = cols(
  participant = col_character(),  # Set 'participant' as character type
  num_correct_trials = col_double(),
  mean_RT = col_double(),
  median_RT = col_double(),
  perc_accuracy = col_double()
))


```


## Digit span tasks combined

```{r}

# Combine data frames using left_join
combined_df <- demo_df %>% 
  left_join(typing_df,  by = "participant") %>%
  left_join(digit_df, by = "participant") %>%
  left_join(vpt_df, by = "participant") %>% 
  left_join(mult_df, by = "participant") %>% 
  filter(!is.na(median_RT))

combined_df <- combined_df %>%
  filter(!is.na(max_span)) %>% 
  pivot_wider(
    names_from = digitspan_type,  # Creates separate columns for each unique value in digitspan_type
    values_from = max_span,  # Specifies the values to be placed in the new columns
    names_prefix = "max_span_"  # Prefix for the new column names
  ) %>% 
  filter(!is.na(max_span_WAIS)) %>% 
 filter(!is.na(max_span_Egner))

combined_df <- combined_df %>%
  mutate(mean_digitspan = (max_span_WAIS + max_span_Egner)/2)


```


# Correlation

```{r}
## Correlation for two digit span tasks
# Convert to numeric
combined_df$max_span_WAIS <- as.numeric(combined_df$max_span_WAIS)
combined_df$max_span_Egner <- as.numeric(combined_df$max_span_Egner)

# Calculate the correlation coefficient and remove missing values
correlation <- cor(combined_df$max_span_WAIS, combined_df$max_span_Egner, use = "pairwise.complete.obs")

# Calculate the p-value for the correlation
correlation_test <- cor.test(combined_df$max_span_WAIS, combined_df$max_span_Egner, method = "pearson")

# Extract the p-value from the test result
p_value <- correlation_test$p.value

# Print or view the p-value
print(p_value)

# Print or view the correlation coefficient
print(correlation)

# Final participants in the digit span task
final_participants <- combined_df %>%
  distinct(participant)

combined_df <- demo_df %>% 
  left_join(typing_df,  by = "participant") %>%
  left_join(digit_df, by = "participant") %>%
  left_join(vpt_df, by = "participant") %>% 
  left_join(mult_df, by = "participant") %>% 
  filter(!is.na(median_RT))

final_participants <- combined_df %>%
  distinct(participant)
```


# Demographics of participants

```{r}

summary_df <-final_participants %>% 
  

```


## Combine date frames

```{r}

interference_score <- readr::read_csv("consolidated_data/stim_inteference_score.csv")


```


```{r}
# add an interference column to mult_df
mult_df <- mult_df %>% 
  left_join(interference_score, by = "problem") 


# Save the combined data frame to a CSV file
write_csv(combined_df, "combined_data.csv")

# Read the CSV file and assign it to combined_rt_df
missing_task_df <- readr::read_csv("combined_data.csv") %>% 
  filter(multiplyCorr == 1) %>% 
  mutate(log10_rt = log10(multiplyRT)) %>%  # Log10 transformed the RT
  group_by(participant) %>%
  summarize_all(~ sum(is.na(.))) %>% 
  filter_at(vars(-participant), any_vars(. != 0))

# Missing task data - participants 685, 749, 783, 947, 971, 979, 987 were excluded
excluded_participants <- missing_task_df$participant

# Load df
combined_df <- read_csv("combined_data.csv") %>% 
  # Exclude participants with missing task data
  filter(!(participant %in% excluded_participants)) %>% 
  # Apply log10 transformation to reaction times
  mutate(log10_rt = log10(multiplyRT))

combined_rt_df <- readr::read_csv("combined_data.csv") %>% 
  # Filter out incorrect trials
  filter(multiplyCorr == 1) %>% 
  # Exclude participants with missing task data
  filter(!(participant %in% excluded_participants)) %>% 
  # Apply log10 transformation to reaction times
  mutate(log10_rt = log10(multiplyRT))

# digit span df
digit_df <- digit_df %>% 
  filter(!(participant %in% excluded_participants)) 

# VPT df
vpt_df <- vpt_df %>% 
  filter(!(participant %in% excluded_participants)) 

# typing df
typing_df <- typing_df %>% 
  filter(!(participant %in% excluded_participants)) 

```

# Descriptive for digit span task
```{r}

# Calculate summary statistics using dplyr
summary_stats <- digit_df %>%
  summarise(
    mean_max_digitspan = mean(max_digitspan),
    median_max_digitspan = median(max_digitspan),
    sd_max_digitspan = sd(max_digitspan)
  )

# Calculate skewness and kurtosis using the psych package
skewness_value <- skew(digit_df$max_digitspan)
kurtosis_value <- kurtosi(digit_df$max_digitspan)

# Create a dataframe for summary statistics
digitspan_summary <- data.frame(
  mean = summary_stats$mean_max_digitspan,
  median = summary_stats$median_max_digitspan,
  sd = summary_stats$sd_max_digitspan,
  skewness = skewness_value,
  kurtosis = kurtosis_value
)
```

# Descriptive for vpt
```{r}

# Calculate summary statistics using dplyr
summary_stats <- vpt_df %>%
  summarise(
    mean_mean_blocks = mean(mean_blocks),
    median_mean_blocks = median(mean_blocks),
    sd_mean_blocks = sd(mean_blocks)
  )

# Calculate skewness and kurtosis using the psych package
skewness_value <- skew(vpt_df$mean_blocks)
kurtosis_value <- kurtosi(vpt_df$mean_blocks)

# Create a dataframe for summary statistics
vpt_summary <- data.frame(
  mean = summary_stats$mean_mean_blocks,
  median = summary_stats$median_mean_blocks,
  sd = summary_stats$sd_mean_blocks,
  skewness = skewness_value,
  kurtosis = kurtosis_value
)
```

# Descriptive for typing
```{r}

# Calculate summary statistics for median_RT using dplyr
median_rt_summary <- typing_df %>%
  summarise(
    mean_median_RT = mean(median_RT),
    median_median_RT = median(median_RT),
    sd_median_RT = sd(median_RT)
  )

# Calculate skewness and kurtosis for median_RT using the psych package
skewness_median_rt <- skew(typing_df$median_RT)
kurtosis_median_rt <- kurtosi(typing_df$median_RT)

# Create a dataframe for median_RT summary statistics
median_rt_summary_df <- data.frame(
  mean = median_rt_summary$mean_median_RT,
  median = median_rt_summary$median_median_RT,
  sd = median_rt_summary$sd_median_RT,
  skewness = skewness_median_rt,
  kurtosis = kurtosis_median_rt
)

```

# Multiplication production task
```{r}

# Calculate summary statistics for multiplication production task using dplyr
combined_acc_summary <- combined_df %>%
  group_by(participant) %>%
  summarise(
    mean_multiplyCorr = mean(multiplyCorr),
    median_multiplyCorr = median(multiplyCorr),
    sd_multiplyCorr = sd(multiplyCorr)
  ) %>%
  ungroup()

# Calculate grand mean of the summary statistics
grand_mean_acc <- combined_acc_summary %>%
  summarise(
    grand_mean_mean = mean(mean_multiplyCorr),
    grand_mean_median = mean(median_multiplyCorr),
    grand_mean_sd = mean(sd_multiplyCorr)
  )

# Calculate skewness and kurtosis using the psych package for grand mean
skewness_value_acc <- skew(combined_acc_summary$mean_multiplyCorr)
kurtosis_value_acc <- kurtosi(combined_acc_summary$mean_multiplyCorr)

# Create a dataframe for summary statistics
multiplication_acc_summary <- data.frame(
  mean = grand_mean_acc$grand_mean_mean,
  median = grand_mean_acc$grand_mean_median,
  sd = grand_mean_acc$grand_mean_sd,
  skewness = skewness_value_acc,
  kurtosis = kurtosis_value_acc
)

# -------------------------------------

# Calculate summary statistics for multiplication production task using dplyr
combined_rt_summary <- combined_rt_df %>%
  group_by(participant) %>%
  summarise(
    mean_multiplyRT = mean(multiplyRT),
    median_multiplyRT = median(multiplyRT),
    sd_multiplyRT = sd(multiplyRT)
  ) %>%
  ungroup()

# Calculate grand mean of the summary statistics
grand_mean_rt <- combined_rt_summary %>%
  summarise(
    grand_mean_mean = mean(mean_multiplyRT),
    grand_mean_median = mean(median_multiplyRT),
    grand_mean_sd = mean(sd_multiplyRT)
  )

# Calculate skewness and kurtosis using the psych package for grand mean
skewness_value_rt <- skew(combined_rt_summary$mean_multiplyRT)
kurtosis_value_rt <- kurtosi(combined_rt_summary$mean_multiplyRT)

# Create a dataframe for summary statistics
multiplication_rt_summary <- data.frame(
  mean = grand_mean_rt$grand_mean_mean,
  median = grand_mean_rt$grand_mean_median,
  sd = grand_mean_rt$grand_mean_sd,
  skewness = skewness_value_rt,
  kurtosis = kurtosis_value_rt
)
```


# Descriptive statistics

```{r}
# Add a task column to each summary dataframe
median_rt_summary_df <- median_rt_summary_df %>% mutate(task = "typing")
vpt_summary <- vpt_summary %>% mutate(task = "vpt")
digitspan_summary <- digitspan_summary %>% mutate(task = "digit span")
multiplication_acc_summary <- multiplication_acc_summary %>% mutate(task = "multiplication - acc")
multiplication_rt_summary <- multiplication_rt_summary %>% mutate(task = "multiplication - rt")

# Combine the three dataframes using bind_rows
final_summary <- bind_rows(
  digitspan_summary,
  vpt_summary,
  median_rt_summary_df,
  multiplication_acc_summary,
  multiplication_rt_summary
  )

# Export the summary dataframe to a CSV file
write.csv(final_summary, "descriptive_summary.csv", row.names = FALSE)
```


## Correlation between digit span, vpt, typing, and multiplication task

```{r}

# Extract relevant columns for each task
digit_span_data <- digit_df %>%
  dplyr::select(participant, max_digitspan)

typing_data <- typing_df %>%
  dplyr::select(participant, median_RT)

vpt_data <- vpt_df %>%
  dplyr::select(participant, mean_blocks)


# ---------------------------------------------------------
# Calculate summary statistics for multiplication production task using dplyr
combined_acc_summary <- combined_df %>%
  group_by(participant) %>%
  summarise(
    mean_multiplyCorr = mean(multiplyCorr),
    median_multiplyCorr = median(multiplyCorr),
    sd_multiplyCorr = sd(multiplyCorr)
  ) %>%
  ungroup()

# Calculate summary statistics for multiplication production task using dplyr
combined_rt_summary <- combined_rt_df %>%
  group_by(participant) %>%
  summarise(
    mean_multiplyRT = mean(multiplyRT),
    median_multiplyRT = median(multiplyRT),
    sd_multiplyRT = sd(multiplyRT)
  ) %>%
  ungroup()

multiplication_rt_data <- combined_rt_summary %>%
  dplyr::select(participant, median_multiplyRT)

multiplication_acc_data <- combined_acc_summary %>%
  dplyr::select(participant, mean_multiplyCorr)

# Combine the data to form a single df
all_corr_data <- full_join(
  digit_span_data,
  typing_data,
  by = "participant"
) %>%
  full_join(vpt_data, by = "participant") %>%
  full_join(multiplication_rt_data, by = "participant") %>%
  full_join(multiplication_acc_data, by = "participant")

```

# Correlation table
```{r}
library(corrr)

# Calculate the correlation matrix
correlation_matrix <- all_corr_data %>%
  dplyr::select(-participant) %>%
  correlate()

# Extract the correlation matrix as a data frame
correlation_df <- correlation_matrix %>%
  shave() %>%
  as.data.frame()

# Calculate the p-values using cor_pmat
p_value_matrix <- correlation_matrix %>%
  shave() %>%
  cor_pmat() %>%
  as.data.frame()

# Correlation matrix

```




## Multiplication production task

### 
